### 2014-10-14 PrivacyProtection-2014-09-09.tex
added bibtex for privacy protection references, still unclear on citation of legal documents

@@

\newcommand{\om}{[...]\xspace} % omission
%\newcommand{\om}{\ldots} % omission

@@

% \section{Histrocial Approaches to Privacy}

@@

\subsection{European Convention on Human Rights}

The \emph{Europen Convetion on Human Rights} \cite{ECHR} was created by the members of the \emph{Council of Europe} (CoE) in 1953 as part of the aftermath of the second world war.
It formulates universal rights of citizens against the state authority.
All member states of the Council of Europe, which includes all EU members, have incorporated this convention into their national law.

Article 8 of the ECHR contains a protection of personal data.

\begin{quote}
(1) Everyone has the right to respect for his private and family life, his home and his correspondence.

(2) There shall be no interference by a public authority with the exercise of this right except such as is in accordance with the law and is necessary in a democratic society in the interests of national security, public safety or the economic well-being of the country, for the prevention of disorder or crime, for the protection of health or morals, or for the protection of the rights and freedoms of others.
\end{quote}

The protection of personal data under Article 8 is not an absolute law but must be considered in relation to other laws such as the freedom of expression (ECHR, Art. 10).
The \emph{European Court of Human Rights} (ECtHR) overseas the implementation of this directive and has adopted a rather broad interpretation of the article.\footnote{\url{http://en.wikipedia.org/wiki/Article_8_of_the_European_Convention_on_Human_Rights}}

In particular State actions of searching a persons home, gathering and storing information in a secret police file, and stopping a prisoner's communication have been ruled to interfere with Article 8.

In the context of the Live+Gov project the ECHR is of limited relevance, since the Live+Gov System is offered by private companies and not directly by the State authority.

\subsection{OECD Guidelines and CoE Convention 108}

The emergence of information technologies that allow automated processing of personal data made more detailed rules for safeguarding the privacy of citizens necessary.
In the light of this developments the OECD  \cite{OECD80} issued Guidelines on Protection of Privacy in 1980 which establish the following basic principles of data protection:
\begin{enumerate}
\item Collection Limitation Principle.
  There should be limits to the collection of personal data and any such data should be obtained \om with the knowledge or consent of the data subject.
\item Data Quality Principle.
  Personal data should be relevant to the purposes for which they are to be used, and,  \om accurate, complete and kept up-to-date.
\item Purpose Specification Principle.
  The purposes for which personal data are collected should be specified \om and the subsequent use limited to the fulfillment of those purposes \om.
\item Use Limitation Principle.
  Personal data should not be disclosed \om except: a) with the consent of the data subject; or b) by the authority of law.
\item Security Safeguards Principle.
  Personal data should be protected by reasonable security safeguards \om.
\item Openness Principle.
  {\om} Means should be readily available of establishing the existence and nature of personal data, and the main purposes of their use, as well as the identity and usual residence of the data controller.
\item Individual Participation Principle.
  An individual should have the right
  \begin{itemize}
  \item to obtain from a data controller, or otherwise, confirmation of whether or not the data controller has data relating to him;
  \item to have communicated to him, data relating to him \om
  \item \om to have the data erased, rectified, completed or amended.
  \end{itemize}
\item Accountability Principle.
  A data controller should be accountable for complying with measures which give effect to the principles stated above.
\end{enumerate}

Where the following definitions are understood.
\begin{itemize}
\item \emph{personal data} means any information relating to an identified or identifiable person (\emph{data subject}).
  An identifiable person is is one who can be identified either directly, by reference to a name or identification number or indirectly,
  by one or more factors which make it possible to find out who the data subject is by conducting further research.
\item a \emph{data controller} means any party who is competent to decide about the contents and use of personal data.
\end{itemize}

The OECD guidelines were adopted by EU law in the Strassbourg Convention 108 of the Council of Europe \cite{CONV108} in 1981.

Relevant for our investigations is in particular, that sensor data collected from mobile devices is considered personal data if the individual can be identified based on the data by any direct or indirect means.
Therefore the above principles apply.

% UPDATE 2013: http://www.oecd.org/sti/ieconomy/privacy.htm

@@

The \emph{Directive 95/46/EC} of 1995 \cite{DIR95} is the fundamental regulation of data protection in the European Union.
It was designed to give further substance to the Convention 108.
In particular it makes the creation of an independent supervisory authority necessary (Art. 28).
As this text is the main legal basis for our investigation we review it in detail.
In Section \ref{sec:ECIMPL} the status of the implementation is discussed in some example cases.

The interpretation and oversight of the Directive lies at the Court of Justice of the European Union (CJEU).
A summary of the most important rulings can be found in Handbook on European data protection law \cite{EU_HANDBOOK_2014}.

The ambition of the Directive is stated at the very beginning.
\begin{quote}
  (Art. 1.1) In Accordance with this Directive the Member States shall protect the fundamental rights and freedoms of natural persons, and in particular their right to privacy with respect to the processing of personal data.
\end{quote}

The Directive follows the definition of personal data, data subject, and data controller of the OECD Guidelines (cf. Art. 2).
Furthermore it introduces the role of a \emph{processor} as a person which processes, by wholly or partly automatic means (Art 3.), personal data on behalf of the controller.

Data collection of intelligence agencies and the police is not restricted by the Directive (Art. 3).

The most important articles are summarized in the following paragraphs.
In order to keep the text more readable we have simplified the formulation by removing the formulation ``Member States shall provide that ...'' from each paragraph.
\begin{itemize}
\item (Art. 6.1.a) Personal data must be processed fairly and lawfully.

[The meaning of the term 'fair processing' is entailed in the following articles.
In particular transparency of processing, information of the data subject, and right to access the data are implied (cf. \cite{EU_HANDBOOK_2014})]

\item (Art. 6.1.b) \emph{Specification of Purpose.} Personal data must be collected for specified, explicit and legitimate purposes and not be processed in a way incompatible with this purposes.

Further processing of data for historical, statistical or scientific purposes shall not be considered incompatible \om.

\item (Art. 6.1.c) Personal data must be adequate, relevant and non-excessive in relation to the purposes \om.

\item (Art. 6.1.d) \emph{Accuracy.} Personal data must be accurate and \om every reasonable step must be taken to ensure that data which is inaccurate or incomplete \om are erased or rectified.

\item (Art. 6.1e) Personal data must be kept in a form which permits identification for no longer than is necessary \om.
% \item (Art. 6.2) It shall be for the controller to ensure that (Art. 6.1) is complied with.
\end{itemize}

In addition to strengthening the OECD Data Quality Principle in Article 6, the Directive requires a strong form of consent of the data subject, before any collection or processing of data can take place.

\begin{itemize}
\item (Art. 7). Personal data may be processed only if:
  \begin{itemize}
    \item [(a)] the data subject has unambiguously given his consent; or
    \item [(..)] [exceptions are made for performance of a contracts with the data subject,  legal obligations of the controller, to protect vital interests of the data subject, the public interest and]
     \item [(f)] the processing is necessary for purposes of legitimate interests pursued by the controller or by a third party to whom the data are disclosed, except where such interests are overridden by the interests \om of the data subject which require protection under Art. 1(1).
  \end{itemize}

\item (Art. 8.1) \emph{Special Categories.} Member Stats shall prohibit the processing of personal data revealing racial or ethnic origin, political opinions, religious or philosophical beliefs, trade-union membership, and the processing of data concerning health or sex life. [Art. 8.2 lists several exceptions, including explicitly explicit consent of the data subject.]
\end{itemize}

The OECD Participation Principle is addressed and extended in the following articles.

\begin{itemize}
\item (Art. 10, 11) The controller \om must provide a data subject \om with the following information:
  \begin{itemize}
    \item [(a)] the identity of the controller \om
    \item [(b)] the purpose of processing for which the data are intended;
    \item [(c)] any further information \om that is necessary to guarantee fair processing in respect of the data subject. [This includes in particular the recipients of the data, the existence of the right to of access and the right to rectify the data.]
  \end{itemize}

\item (Art. 12) \emph{Right to Access.} Every data subject has the right to obtain from the controller
\begin{itemize}
  \item (a) \om confirmation as to whether or not data relating to him are being processed \om, communication to him \om the data undergoing processing and any available information as to their source, knowledge of the logic involved in any automatic processing \om;
  \item (b) the rectification, erasure or blocking of data the procession of which does not comply with the provisions of this Directive \om;
  \item (c) notification to third parties to whom the data have been disclosed \om.
\end{itemize}

%\item Confidentiality of Processing (Art. 16).
%Any person under the authority of the controller or of the processor \om who has access to personal data must not process them except on instructions from the controller, unless he is required to do so by law.
\end{itemize}

A similar form of the Security Safeguard Principle is contained in Article 17.

\begin{itemize}
\item (Art. 17) \emph{Security of processing.}
The controller must implement appropriate technical and organizational measures to protect personal data against accidental loss, alteration, unauthorized disclosure or access \om and against all other forms of unlawful processing.

[The level of security has to be balanced against the risk of processing.]

% Having regard to the state of the art and the cost of their implementation, such measures shall ensure a level of security appropriate to the risk represented by the processing and the nature of the data to be processed.
\end{itemize}

Furthermore the Directive demands all data processing to be reported to a supervisory authority.
This obligation can be lifted if an internal data protection official is appointed, who is responsible in particular for keeping a processing register, that has to be made available on request.

\begin{itemize}
\item Notification (Art. 18. 1).
The controller \om must notify the supervisory authority referred to in Article 28 before carrying out any wholly or partly automatic processing operation \om.

\item (Art. 18.2).
Simplification or exemption from notification may be provided only under the following conditions: \om where the controller \om appoints a personal data protection official, responsible in particular:
\begin{itemize}
\item for ensuring in an independent manner the internal application of the national provisions taken pursuant to this directive.
\item for keeping the register of processing operations carried out by the controller, containing the items of information referred to in Article 21 (2)
\end{itemize}
thereby ensuring that the rights and freedoms of the data subjects are unlikely to be adversely affected by the processing operations.

\item (Art. 19.1)
The information to be given in the notification shall include at least:
\begin{itemize}
  \item [(a)] the name and address of the controller \om;
  \item [(b)] the purpose \om of processing;
  \item [(c)] a description of the category \om of the data \om;
  \item [(d)] the recipients \om to whom the data might be disclosed;
  \item [(e)] proposed transfer of data to third countries;
  \item [(f)] a general description allowing a preliminary assessment to be made of the appropriateness of the measures taken pursuant of Article 17 to ensure the security of processing;
\end{itemize}

\item (Art. 21.2) \emph{Publication}.
A register of processing information \om shall be kept at the supervisory authority.
The register shall contain at least the information listed in Article 19.1 (a) to (e).

The register may be inspected by any person.

\item (Art. 21.3). In relation to processing operations not subject to notification, that controllers \om make available at least the information referred to in Article 19.1 (a) to (e) in an appropriate form to any person on request.

\item (Art. 28) \emph{Supervisory Authority.}
Each Member State shall provide that one or more public authorities are responsible for monitoring the application \om of this Directive.

These authorities shall act in complete independence in exercising this functions entrusted to them.
\end{itemize}

Data transfer to third countries outside of the EU requires those countries to have an adequate level of data protection.
\begin{itemize}
\item (Art. 25.1) \emph{Transfer to third countries.}
The transfer to a third country of personal data \om may take place only if \om the third country in question ensures an adequate level of protection.
\end{itemize}
The European Commission can decide whether a third country ensures an adequate level of protection (Art 25.6).
The USA, is not considered to do so.
Data exchange between the US and EU is possible under the Safe Harbor regulation \cite{SAFE_HABOR}.
Further Exceptions are provided by the Passenger Name Record Agreement \cite{PNR}.

\subsection{Implementation of the Data Protection Directive}\label{sec:ECIMPL}

\textbf{Germany}.
Germany implements Directive 95/46 with the \emph{Bundesdatenschutzgesetz (BDSG)}
of 2001. However, Germany has violated the directive in two points:
\begin{enumerate}
\item The BDSG has become effective three years too late, thus the EC filed a treaty violation proceeding against Germany.
\item The BDSG does not implement independent supervisory authorities.
The Bundesdatenschutzbeauftragter is subordinate to the Ministry of Interior.
Although he is not subject to technical oversight (\emph{Fachaufsicht}), he is subject to staff supervision by the government (Rechtsaufsicht, Dienstaufsicht) and budget oversight by the ministry.
In March 2010 Germany was found guilty of violation of Directive 95/46 by the ECJ.
\end{enumerate}

The states of Germany have their own implementation of Directive 95/46 (\emph{Landesdatenschutzgesetze}).
Federal public authorities are only bound to their federal law.
Churches are not subject the BDSG.

\textbf{United Kingdom.}
The UK implements Directive 95/46 with the \emph{Data Protection Act 1998} (DPA).

The act is known for its high complexity: a manual record of phone numbers for business purposes could be hold subject to the DPA.
Although the act seems to fully cover the directive.
Even higher restriction apply for \emph{``sensitive personal data''} (race, ethnicity, politics, religion, trade union status, health, sex life or criminal record), i.e. consent must be given freely and has to be explicit.

The Act's definition of''personal data" covers any data that can be used to identify a living individual.
Anonymised or aggregated data is not regulated by the Act, providing the anonymisation or aggregation has not been done in a reversible way.
Individuals can be identified by various means including their name and address, telephone number or Email address.
The Act applies only to data which is held, or intended to be held, on computers (`equipment operating automatically in response to instructions given for that purpose'), or held in a `relevant filing system'.

In some cases even a paper address book can be classified as a `relevant filing system', for example diaries used to support commercial activities such as a salesperson's diary.

The Freedom of Information Act 2000 modified the act for public bodies and authorities, and the Durant case modified the interpretation of the act by providing case law and precedent.

The Data Protection Act creates rights for those who have their data stored, and responsibilities for those who store, process or transmit such data.
The person who has their data processed has the right to:
\begin{itemize}
\item View the data an organization holds on them.
A `subject access request' can be obtained for a nominal fee.
As of January 2014, the maximum fee is \pounds2 for requests to credit reference agencies, \pounds50 for health and educational request, and \pounds10 per individual otherwise.
\item Request that incorrect information be corrected.
If the company ignores the request, a court can order the data to be corrected or destroyed, and in some cases compensation can be awarded.
\item Require that data is not used in any way that may potentially cause damage or distress.
\item Require that their data is not used for direct marketing. \cite{3}
\end{itemize}

@@

\subsection{EU Charter of Fundamental Rights}

The Treaty of Lisbon of 2007, which layed down the fundamental functioning of the European Union
includes the \emph{Charter of Fundamental Rights of the Europen Union} \cite{EUFR2010}.
This Charta summarizes the full range of civil, political and economic rights of EU citizens and contains the following two articles that safeguard the privacy of the citizen.
\begin{quote}
  Article 7. Respect for private and family life.\\
  Everyone has the right to respect for his or her private and family life, home and communications.

  Article 8. Protection of personal data.
  \begin{enumerate}
    \item [(1)] Everyone has the right to the protection of personal data concerning him or her.
    \item [(2)] Such data must be processed fairly for specified purposes and on the basis of the consent of the person concerned or some other legitimate basis laid down by law.
      Everyone has the right of access to data which has been collected concerning him or her, and the right to have it rectified.
    \item [(3)] Compliance with these rules shall be subject to control by an independent authority.
  \end{enumerate}
\end{quote}

Although the Charter does not extend the pre-exiting Directive 95/46/EC, it amplifies the importance of this privacy protection as a fundamental right.

@@

This type is also concerned with the ``protection against disclosure of personal matters'' through behaviour, however Clarke's distinction between ``casual observation \om systematic recording and storage of information about those activities'' is lifted.

@@

The \textbf{Privacy of The Person} is generally concerned with one could
best understand as \emph{Biometric Privacy}. Friedewald et al. paraphrase
it as \emph{``\om the right to keep body functions and body
characteristics \om private''}. \textbf{Accelerometer},
\textbf{Rotation Vector} and \textbf{Gyroscope} measure the physical
movement of the mobile device on all three axes. If the mobile device
is carried \emph{``normally''} its safe to say that those sensors also
measure the moments of its carrier. So his privacy is infringed
regarding biometric behaviour, as it is captured automatically.
(\textbf{Note:} This is not to be confused with the \textbf{Privacy of
Behaviour and Action} which is used for the social aspects of behaviour
action, e.g.~praying, sexual habits or political activities.)

>>>>

% test citing
\cite{CFried:Privacy}
\cite{WarrenBrandeis:RightToPrivacy}
\cite{sep-privacy}
\cite{7ToP}
\cite{Grimm:ItSecRefModel}
\cite{ItSecGlossary}
\cite{SociotechnicalArchitectureForOnlinePrivacy}
\cite{EUDataProtectionDirective}
\cite{EUConventionOnHumanRights}
\cite{Bundesdatenschutzgesetz}
\cite{Gutwirth:DefiningPrivacy}
\cite{UKDataProtectionAct}

@@

@@

\section{Histrocial Approaches to Privacy}

\subsection{Aristotle}

Public Sphere vs. Private Sphere of politics. This distinction is concerned with the politic (public) life of politicians in early democracies opposed by their domestic (private) life.

\subsection{Warren and Brandeis}

The Right to Privacy, 1890: Establish the term Informational Privacy. They argue innovations like photography and a growing press create the need for a new right to privacy, "the right to be left alone".

The term informational privacy as conceptual foundation of privacy originates from the juristic discussion The Right to Privacy (1890) by Warren and Brandeis.
The arise of new technologies like photography combined with a growing press created an increased publicity.
Warren and Brandeis were worried that existing law like castle doctrines or libel and slander could not suffice.
They felt a need for a new right to "protect the extent to which one's thoughts, sentiments, and emotions could be shared with others" \cite{1}.
This right would be "the right to be left alone".

The right to "protect the extent to which one's thoughts, sentiments, and emotions could be shared with others" by Warren and Brandeis was later extend to cover arbitrary information and is now kown as the privacy concept "... the control we have over information about ourselves ..." \cite{2}.

\subsection{Fried}

C. Freid, An Anatomy of Values, Harvard Univ. Press, 1970

Defines privacy as the ability to control information about oneself. He argues that limited knowledge does not automatically create privacy.

@@

\subsection{European Convention on Human Rights - Article 8}

``(1) Everyone has the right to respect for his private and family life, his home and his correspondence. (2) There shall be no interference by a public authority with the exercise of this right except such as is in accordance with the law and is necessary in a democratic society in the interests of national security, public safety or the economic well-being of the country, for the prevention of disorder or crime, for the protection of health or morals, or for the protection of the rights and freedoms of others.'' \cite{2}

``The European Court of Human has given this article a very broad interpretation in its jurisprudence.'' \cite{2}

@@

\emph{Directive 95/46/EC}

EU directives in general do not hold any direct legal binding for EU citizens.
Member states have to create their own legal implementation.

``The directive regulates the processing of personal data regardless of whether such processing is automated or not.'' \cite{1} where \emph{personal data} is ``any information relating to an identified or identifiable natural person [...] (art. 2a)'' and \emph{processing} is ``any operation or set of operations which is performed upon personal data [...] (art. 2b)'' \cite{1}

The ``identified or identifiable natural person'' is called \emph{data subject}.

The sole responsibility for compliance is held by so called \textbf{controllers}.
A controller is an actor ``which alone or jointly with others determines the purposes and means of the processing of personal data; (art 2d)''
\cite{1}

This rule is applicable ``whenever the controller uses equipment situated within the EU''\cite{1}.
This will hold for every e-commerce provider \textbf{inside or outside} of the EU because the customer's computer is \emph{situated within the EU} anyways.

``Personal data should not be processed at all, except when certain conditions are met'' \cite{1}.
These conditions incorporate the seven OECD principals and are further categorized in the categories:

\begin{itemize}
\item
\textbf{Transparency}: ``The data subject has the right to be informed when his personal data is being processed [...] (art. 10 and 11)'' \cite{1} and he has the right how and by whom it is processed.
Additionally other specified requirements have to be met, i.e.~giving consent.
\item
\textbf{Legitimate purpose}: ``Personal data can only be processed for specified explicit and legitimate purposes and may not be processes further in a way incompatible with those purposes (art. 6b)''
\cite{1}.
\item
\textbf{Proportionality}: ``Personal data may be processed only insofar as it is adequate [...] in relation to [its] purposes [...]; every reasonable step must be taken to ensure that data which are inaccurate or incomplete [...] are erased or rectified (art. 6)'' \cite{1}
\end{itemize}

(\textbf{Note:} The principles incorporated by the EU directive seem to closely map to \emph{the seven principles for user privacy control}.)

Member states have to create independent \textbf{supervisory authorities} for personal data processing where controllers have to register before they start to process any data.
The record has to be stored in public register \cite{1}.

Data transfer to third countries outside of the EU requires those countries to have a similar data protection level. Exceptions exists:

\begin{itemize}
\item
  Save Harbor
\item
  Passenger Name Record Agreement
\end{itemize}

all between the US and the EU, although the US has no comparable law of data protection. 
The US Supreme Court regards privacy as \textbf{implicit} right granted with the First Amendment. However, European law states privacy as \textbf{explicit} right (ECHR Art. 8).
Additionally the US government follows a doctrine which favors the economy to implement self-regulation \cite{1}.

\subsection{Implementation of the Data Protection Directive}

\subsubsection*{Germany}

Germany implements Directive 95/46 with the \emph{Federal Data Protection Act} from 2001 known as \emph{Bundesdatenschutzgesetz (BDSG)} \cite{1}\cite{2}\cite{6}.
However, Germany has violated the directive in two points:

\begin{enumerate}
\item
The BDSG has become effective three years too late, thus the EC filed a treaty violation proceeding against Germany \cite{6}.

\item
The BDSG does not implement \textbf{independent} supervisory authorities. 
The \emph{Bundesdatenschutzbeauftragter} is subordinate to the Ministry of Interior. 
Although he is not subject to technical oversight (\emph{Fachaufsicht}), he is subject to staff supervision by the government (\emph{Rechtsaufsicht} durch die Bundesregierung und  \emph{Dienstaufsicht} durch das Innenministerium) and budget oversight (including approval of employees) by the ministry \cite{7}. 
Thus the EC filed a treaty violation proceeding against Germany, again \cite{6}. 
In March 2010 Germany was found guilty of violation of Directive 95/46 by the ECJ \cite{6}.
\end{enumerate}

Germany still violates point 2!

The states of Germany have their own implementation of Directive 95/46 (\emph{Landesdatenschutzgesetze}).
Federal public authorities are only bound to their federal law. \cite{2}

Churches are not subject the BDSG. \cite{2}

\item
\textbf{Roman Catholic Church:}
\href{http://de.wikipedia.org/wiki/Anordnung_\%C3\%BCber_den_kirchlichen_Datenschutz}{Anordnung  über den kirchlichen Datenschutz (KDO)}

\item
\textbf{German Protestant Churches (Synode der Evangelischen Kirche in Deutschland):}
\href{http://de.wikipedia.org/wiki/Datenschutzgesetz_der_Evangelischen_Kirche_in_Deutschland}{Kirchengesetz  über den Datenschutz der Evangelischen Kirche in Deutschland}
\end{itemize}

Data subjects have the following rights.

\begin{itemize}
\item
Right to disclosure of whether data about them is stored and processed, which data is stored and processed and the data sources.
\item
Right to correction of false data.

\item
Right to file complaints to the supervisory authority.

\item
Right to have data deleted or blocked, but controllers can prohibit deletion in favour of blocking.

\item
Right to decline third party access to the data.
\end{itemize}

\subsubsection*{United Kingdom}

The UK implements Directive 95/46 with the \emph{Data Protection Act 1998 (DPA)} \cite{1}\cite{3}.

The act is known for its high complexity: a manual record of phone numbers for business purposes could be hold subject to the DPA \cite{3}.
Although the act seems to fully cover the directive.
Even higher restriction apply for \emph{``sensitive personal data''} (race, ethnicity, politics, religion, trade union status, health, sex life or criminal record), i.e. \textbf{consent} must be given freely and has to be explicit. \cite{3}

\emph{``The Act's definition of''personal data" covers any data that can be used to identify a living individual. Anonymised or aggregated data is not regulated by the Act, providing the anonymisation or aggregation has not been done in a reversible way.
Individuals can be identified by various means including their name and address, telephone number or
Email address.
The Act applies only to data which is held, or intended to be held, on computers (`equipment operating automatically in response to instructions given for that purpose'), or held in a `relevant filing system'.}

\emph{In some cases even a paper address book can be classified as a `relevant filing system', for example diaries used to support commercial activities such as a salesperson's diary.}

\emph{The Freedom of Information Act 2000 modified the act for public bodies and authorities, and the Durant case modified the interpretation of the act by providing case law and precedent.}

\emph{The Data Protection Act creates rights for those who have their data stored, and responsibilities for those who store, process or transmit such data.
The person who has their data processed has the right to:}

\item
View the data an organization holds on them.
A `subject access request' can be obtained for a nominal fee. As of January 2014, the maximum fee is \pounds2 for requests to credit reference agencies, \pounds50 for health and educational request, and \pounds10 per individual otherwise.

\item
Request that incorrect information be corrected.
If the company ignores the request, a court can order the data to be corrected or destroyed, and in some cases compensation can be awarded.

\item
Require that data is not used in any way that may potentially cause damage or distress.

\item
Require that their data is not used for direct marketing. \cite{3}
\end{itemize}

@@

@@

This type is also concerned with the ``protection against disclosure of personal matters'' through behaviour, however Clarke's distinction between ``casual observation [...] systematic recording and storage of information about those activities'' is lifted.

@@

The \textbf{Privacy of The Person} is generally concerned with one could
best understand as \emph{Biometric Privacy}. Friedewald et al. paraphrase
it as \emph{``[...] the right to keep body functions and body
characteristics [...] private''}. \textbf{Accelerometer},
\textbf{Rotation Vector} and \textbf{Gyroscope} measure the physical
movement of the mobile device on all three axes. If the mobile device
is carried \emph{``normally''} its safe to say that those sensors also
measure the moments of its carrier. So his privacy is infringed
regarding biometric behaviour, as it is captured automatically.
(\textbf{Note:} This is not to be confused with the \textbf{Privacy of
Behaviour and Action} which is used for the social aspects of behaviour
action, e.g.~praying, sexual habits or political activities.)
